---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visage

<!-- badges: start -->
<!-- badges: end -->

The goal of visage is to provide an lightweight OOP system and a set of tools for running visual inference experiments.

## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TengMCing/visage")
```

```{r}
library(visage)
```

## Usage of visual inference cubic linear model class

### Define a model

A visual inference cubic linear model is defined as

```{r}
CUBIC_MODEL$formula
```

where `x` and `z` are random variables, `a`, `b` and `c` are parameters and `e` is the random error term.

To define the model, `cubic_model()` needs to be used. `x` and `z` by default are assumed to follow a uniform distribution and `e` by default is assumed to follow a normal distribution with standard deviation equals to `sigma`.

```{r}
mod <- cubic_model(a = 1, b = 1, c = 1, sigma = 1)
mod
```

To specify the different distribution for `x` and `z` and `e`, you can follow this example:

```{r}
# Normal distribution
x <- rand_normal(mu = 0, sigma = 1)

# Uniform distribution
z <- rand_uniform(a = -1, b = 1)

# Log-normal distribution
e <- rand_lognormal(mu = 0, sigma = 10)

mod2 <- cubic_model(a = 1, b = 1, c = 1, x = x, z = z, e = e)

mod2
```


### Generate random values from the model

To generate random values from the model, method `gen()` needs to be used. It will return a data frame containing values of `y`, `x`, `z`, `e`, `.resid` and `.fitted`.

```{r}
mod$gen(n = 5)
```

Residuals and fitted values are obtained by fitting the following null model:

```{r}
CUBIC_MODEL$null_formula
```

### Make a residual plot

```{r}
mod$plot(mod$gen(100), "resid")
```

### Generate data for a lineup

```{r}
mod$gen_lineup(n = 10, k = 2)
```

### Plot a lineup

```{r}
mod$plot_lineup(mod$gen_lineup(100))
```


## Basic usgae of `visage` OOP system

### Define a new class

A class can be defined with the `new_class` function. All positional arguments are for specifying parent classes, `BASE` is the base object class provided by the package, you don't need to manually specify it. But if you would like to have advanced behaviour, you can try to implement your own `object` class.  

Class name is mandatory and should be unique. 

```{r}
# You don't actually need to specify BASE here. This is only for demonstration.
DEMO <- new_class(BASE, class_name = "DEMO")
DEMO
```

The object is an environment containing some useful attributes and methods.

- `OBJECT$..type..` gives the current class name.
- `OBJECT$..class..` gives the current class name and parent class names.

```{r}
DEMO$..type..
DEMO$..class..
```

- `OBJECT$..dict..()` returns all names of attribute and method of the object.
- `OBJECT$..methods..()` returns all names of method of the object


```{r}
DEMO$..dict..()
DEMO$..methods..()
```

- `OBJECT$..str..()` returns a string representation of the object, which will be used by the S3 `print()` method. This method usually needs to be overridden in subclass to give short summary of the object. 

```{r}
DEMO$..str..()
```


### Register a method for the class

Methods can be registered by using `register_method()`. The first argument is the object you want to bind the function to, the rest of the positional arguments are for specifying method names and functions. The syntax is `method_name = function`.

You can choose to write inline function or pass pre-defined function. The associative environment of the function doesn't matter, it will be modified by the `register_method()` function. 

```{r}
pre_defined_fn <- function() 1 + 2

register_method(DEMO, inline_fn = function() 1 + 1, pre_defined_fn = pre_defined_fn)

DEMO$inline_fn()
DEMO$pre_defined_fn()
```

For method that needs to access the object itself, just simply use `self` in your method. It is an reference to the object. 

```{r}
DEMO$val <- 5

register_method(DEMO, get_val = function() self$val)

DEMO$get_val()
```

### Override the `..init..()` method

`..init..()` method is for instance initialization. To override the `..init..()` method, you need to use the `register_method()` to register it again.

```{r}
init <- function(first_name, employee_id) {
  self$first_name <- first_name
  self$employee_id <- employee_id
}

register_method(DEMO, ..init.. = init)
```

Now the class requires two two arguments `first_name` and `employee_id` to initialize the instance.

### Build an instance

To new and initialize an instance, you need to use the `instantiation()` method. The output will show it is an object.

```{r}
mike <- DEMO$instantiation("Mike", 25)
mike
```

`first_name` and `employee_id` are stored in the object because of the `..init..()` method.

```{r}
mike$first_name
mike$employee_id
```


### A complete workflow

It is recommend to write your class definition in a function to make debugging easier. The following example new a class `DEMO_2`, defines its own `..init..()` method, defines a `get_email()` function for retrieving the email address, defines its own `..str..()` method such that when we print the object, it will provide us with a nicely formatted summary.

`use_method` is used to run methods from other classes, which in this case, the `..str..()` method from the `BASE` class.

```{r}
class_DEMO_2 <- function(env = new.env(parent = parent.frame())) {
  
  new_class(env = env, class_name = "DEMO_2")
  
  init_ <- function(first_name, employee_id) {
    self$first_name <- first_name
    self$employee_id <- employee_id
  }
  
  get_email_ <- function() {
    paste0(self$first_name, "_", self$employee_id, "@company.com")
  }
  
  str_ <- function() {
    paste(use_method(self, BASE$..str..)(), 
          paste("Name:", self$first_name,
                "\nEmployee ID:", self$employee_id,
                "\nEmail:", self$get_email()), 
          sep = "\n")
  }
  
  register_method(env,
                  ..init.. = init_,
                  get_email = get_email_,
                  ..str.. = str_)
  
  return(env)
}
```


```{r}
DEMO_2 <- class_DEMO_2()
mike <- DEMO_2$instantiation("Mike", 25)
mike$get_email()
```

```{r}
mike$..str..()
mike
```
