---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visage

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/TengMCing/visage/branch/master/graph/badge.svg)](https://app.codecov.io/gh/TengMCing/visage?branch=master)
<!-- badges: end -->

The goal of visage is to provide an lightweight OOP system and a set of tools for running visual inference experiments.

## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TengMCing/visage")
```

```{r}
library(visage)
library(bandicoot)
```


## Table of Content

1. [Visual inference cubic linear model class `CUBIC_MODEL`](#1-visual-inference-cubic-linear-model-class-cubic_model)
2. [Visual inference heteroskedasticity linear model class `HETER_MODEL`](#2-visual-inference-heteroskedasticity-linear-model-class-heter_model)


## 1. Visual inference cubic linear model class `CUBIC_MODEL`

**1.1. Define a model**

A visual inference cubic linear model is defined as:

```{r}
CUBIC_MODEL$formula
```

where `x` and `z` are random variables, `a`, `b` and `c` are parameters and `e` is the random error term.

`cubic_model()` is a cubic model class constructor, which takes arguments `a`, `b`, `c`, `sigma`, `x` and `z`, where the first four are numeric values defined above, and `x` and `z` are random variable instances created by the random variable abstract base class constructor `rand_var()`. If we would like `x` and `z` to be uniform random variables ranged from −1 to 1, it can be achieved by using the uniform random variable class constructor `rand_uniform()`. It only takes two arguments which are the lower bound and the upper bound of the support.

```{r}
mod <- cubic_model(a = -3, b = -3, c = 1, sigma = 1,
                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
mod
```


**1.2. Generate random values from the model**

An instance of cubic model class contains methods of simulating data and making residual plot. Method `gen()` returns a data frame containing realizations of `x`, `z`, `y` and `e` simulated from the model. The number of realizations depends on the integer argument `n`. In addition, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.

The null model is defined as:

```{r}
CUBIC_MODEL$null_formula
```

Generates five realizations from the model:

```{r}
mod$gen(n = 5)
```

**1.3. Make a residual plot**

Method `plot()` produce a `ggplot` object. It takes a data frame containing columns `.resid` and `.fitted` as input, along with a character argument type indicating the type of the data plot, and other aesthetic arguments such as size and alpha to control the appearance of the plot.

```{r}
mod$plot(mod$gen(n = 100), type = "resid", size = 1)
```

**1.4. Make a lineup**

A lineup is a matrix of residual plots which can be produced by using the methods `gen_lineup()` and `plot_lineup()`. Method `gen_lineup()` takes the number of realizations `n` and the number of plots in a lineup `k` as inputs. Other than that, the method `plot_lineup()` has the same user interface as `plot()`.

Generate a lineup consists of 2 plots and 8 realizations:

```{r}
mod$gen_lineup(n = 8, k = 2)
```


Plot a lineup consists of 20 plots and 300 realizations: 

```{r}
mod$plot_lineup(mod$gen_lineup(n = 300, k = 20), type = "resid", alpha = 0.6)
```

## 2. Visual inference heteroskedasticity linear model class `HETER_MODEL`

A visual inference heteroskedasticity linear model is defined as:

```{r}
HETER_MODEL$formula
```

where `x` is random variables, `a`, `b` are parameters and `e` is the random error term.

Similar to the cubic model, the heteroskedasticity model could be built by the heteroskedasticity model class constructor `heter_model()`. This function takes three arguments as inputs, which are `a`, `b` and `x`. `a` and `b` are numeric parameters defined above. `x` needs to be a random variable object.

```{r}
mod <- heter_model(a = 0, b = 16, x = rand_uniform(-1, 1))

mod
```


Since both the cubic model class `CUBIC_MODEL` and the heteroskedasticity model class `HETER_MODEL` are inherited from the visual inference model class `VI_MODEL`, heteroskedasticity model object can be used in a similar way as cubic model object. The following codes give examples of the use of the object.

```{r}
mod$gen(n = 5)
```

```{r}
mod$plot(mod$gen(n = 300), type = "resid", size = 1)
```

```{r}
mod$plot_lineup(mod$gen_lineup(n = 300), alpha = 0.6)
```
