---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visage

<!-- badges: start -->
<!-- badges: end -->

The goal of visage is to provide an lightweight OOP system and a set of tools for running visual inference experiments.

## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TengMCing/visage")
```
## Basic examples for usgae of `visage` OOP system

```{r}
library(visage)
```

### Define a new class

A class can be defined with the `new_class` function. All positional arguments are for specifying parent classes, `BASE` is the base object class provided by the package, so you almost always need to include it in . If you would like to have advanced behaviour, you can try to implement your own `object` class.  

Class name is mandatory. 

```{r}
DEMO <- new_class(BASE, class_name = "DEMO")
DEMO
```

