---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visage

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/TengMCing/visage/branch/master/graph/badge.svg)](https://app.codecov.io/gh/TengMCing/visage?branch=master)
<!-- badges: end -->

The goal of visage is to provide a set of tools for running visual inference experiments on linear regression models.

## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_url("https://github.com/TengMCing/visage/raw/master/built/visage_0.1.1.tar.gz")
```

```{r}
library(visage)
```


## Table of Content

1. [Visual inference cubic linear model class `CUBIC_MODEL`](#1-visual-inference-cubic-linear-model-class-cubic_model)
2. [Visual inference heteroskedasticity linear model class `HETER_MODEL`](#2-visual-inference-heteroskedasticity-linear-model-class-heter_model)
3. [Closed form expression `CLOSED_FORM`](#3-closed-form-expression-closed_form)


## 1. Visual inference cubic linear model class `CUBIC_MODEL`

**1.1. Define a model**

A visual inference cubic linear model is defined as:

```{r}
CUBIC_MODEL$formula
```

where `x` and `z` are random variables, `a`, `b` and `c` are parameters and `e` is the random error term.

`cubic_model()` is a cubic model class constructor, which takes arguments `a`, `b`, `c`, `sigma`, `x` and `z`, where the first four are numeric values defined above, and `x` and `z` are random variable instances created by the random variable abstract base class constructor `rand_var()`. If we would like `x` and `z` to be uniform random variables ranged from −1 to 1, it can be achieved by using the uniform random variable class constructor `rand_uniform()`. It only takes two arguments which are the lower bound and the upper bound of the support.

```{r}
mod <- cubic_model(a = -3, b = -3, c = 1, sigma = 1,
                   x = rand_uniform(-1, 1), z = rand_uniform(-1, 1))
mod
```


**1.2. Generate random values from the model**

An instance of cubic model class contains methods of simulating data and making residual plot. Method `gen()` returns a data frame containing realizations of `x`, `z`, `y` and `e` simulated from the model. The number of realizations depends on the integer argument `n`. In addition, a null model will be fitted using the simulated data and residuals and fitted values will be included in the returned data frame.

The null model is defined as:

```{r}
CUBIC_MODEL$null_formula
```

Generates five realizations from the model:

```{r}
mod$gen(n = 5)
```

**1.3. Make a residual plot**

Method `plot()` produce a `ggplot` object. It takes a data frame containing columns `.resid` and `.fitted` as input, along with a character argument type indicating the type of the data plot, and other aesthetic arguments such as size and alpha to control the appearance of the plot.

```{r}
mod$plot(mod$gen(n = 100), type = "resid", size = 1)
```

**1.4. Make a lineup**

A lineup is a matrix of residual plots which can be produced by using the methods `gen_lineup()` and `plot_lineup()`. Method `gen_lineup()` takes the number of realizations `n` and the number of plots in a lineup `k` as inputs. Other than that, the method `plot_lineup()` has the same user interface as `plot()`.

Generate a lineup consists of 2 plots and 8 realizations:

```{r}
mod$gen_lineup(n = 8, k = 2)
```


Plot a lineup consists of 20 plots and 300 realizations: 

```{r}
mod$plot_lineup(mod$gen_lineup(n = 300, k = 20), type = "resid", alpha = 0.6)
```

## 2. Visual inference heteroskedasticity linear model class `HETER_MODEL`

A visual inference heteroskedasticity linear model is defined as:

```{r}
HETER_MODEL$formula
```

where `x` is random variables, `a`, `b` are parameters and `e` is the random error term.

Similar to the cubic model, the heteroskedasticity model could be built by the heteroskedasticity model class constructor `heter_model()`. This function takes three arguments as inputs, which are `a`, `b` and `x`. `a` and `b` are numeric parameters defined above. `x` needs to be a random variable object.

```{r}
mod <- heter_model(a = 0, b = 16, x = rand_uniform(-1, 1))

mod
```


Since both the cubic model class `CUBIC_MODEL` and the heteroskedasticity model class `HETER_MODEL` are inherited from the visual inference model class `VI_MODEL`, heteroskedasticity model object can be used in a similar way as cubic model object. The following codes give examples of the use of the object.

```{r}
mod$gen(n = 5)
```

```{r}
mod$plot(mod$gen(n = 300), type = "resid", size = 1)
```

```{r}
mod$plot_lineup(mod$gen_lineup(n = 300), alpha = 0.6)
```


## 3. Closed form expression `CLOSED_FORM`

**3.1. Define and evaluate a closed form expression**

"Closed form expression" is defined as a special type of `R` expression that can replace random variables in the expression with their realizations before evaluation. This may sound confusing, so let's look at some examples.

`closed_form()` is the closed form expression constructor, which takes only one arguments `expr` required to be a formula. Considering the simplest case first, where the expression contains zero random variables, such as `1 + 3`. It can be evaluated using the method `compute()` directly as a typical `R` expression.

```{r}
cf <- closed_form(~1 + 3)
cf
cf$compute()
```

However, if the closed form expression contains any random variables, it can not be evaluated directly since the values of random variables are unknown. In the following example, `x` is a uniform random variable ranged from -1 to 1, and `x + 1` is an expression that can not be evaluated by method `compute()`.

```{r}
x <- rand_uniform(-1, 1)
cf <- closed_form(~1 + x)
```

In this case, the `gen()` method will try to generate `n` realizations from random variables, then evaluate the expression. In addition, if `rhs_val = TRUE`, realizations generated from random variables will be returned.

```{r}
cf$gen(n = 5, rhs_val = TRUE)
```

**3.2. Nested closed form expression**

Nested closed form expression is useful when a hierarchical structure present in the expression.

Consider the following statistical model:

`y = x + e`, where `x` is a uniform random variable ranged from -1 to 1, and `e = 1 + x^2 * z`, where `z` is a standard normal random variable.

We can define this model using the following code:

```{r}
z <- rand_normal(mu = 0, sigma = 1)
x <- rand_uniform(-1, 1)
e <- closed_form(~1 + x^2 * z)
y <- closed_form(~x + e)
y
```

Note that `y` depends on `e` and `e` itself is a closed form expression. Both `y` and `e` depend on `x`. The method `gen()` will return values of `x`, `z` and `e` if `rhs_val = TRUE`.

```{r}
y$gen(n = 5, rhs_val = TRUE)
```

**3.3. Turn `gen()` result into dataframe via method `as_dataframe()`**

Results of `gen()` is either a vector or a list, which can be converted into a dataframe via method `as_dataframe()`. The `lhs` argument is for naming the left hand side result of the closed form expression.

```{r}
y$as_dataframe(y$gen(n = 5, rhs_val = TRUE), lhs = "y")
```


**3.4. Current limitiation of `CLOSED_FORM`**

For `closed_form()`, any simple expressions can be provided, as long as all the symbols exist in the current environment. Note that this function tries to evaluate **ALL** the atomic symbols in the expression during initialization, and store the values in the object. Hence, calls like `a$b` will also be decomposed as `$`, `a` and `b`, where `b` will be interpreted as a variable `b` exists in the current environment. Therefore, use `~a[["b"]]` instead of `~a$b`. And pre-define function like `myfun = function() 1`, then use it in the expression `~myfun()`.
